Fossil-calibrated phylogeny of Onagraceae: patterns of floral and chromosome evolution

### 2014-04-27 manuscript.tex
much of methods section, reformatting throughout

\usepackage{lineno,hyperref}
\modulolinenumbers[5]

@@

%%%%%%%%%%%%%%%%%%%%%%%
%% Elsevier bibliography styles
%%%%%%%%%%%%%%%%%%%%%%%

%% Harvard
\bibliographystyle{model2-names}\biboptions{authoryear}

%% `Elsevier LaTeX' style
%\bibliographystyle{elsarticle-num}

%%%%%%%%%%%%%%%%%%%%%%%

@@

\begin{abstract}
This template helps you to create a properly formatted \LaTeX\ manuscript.
\end{abstract}

%\begin{keyword}
%\texttt{elsarticle.cls}\sep \LaTeX\sep Elsevier \sep template
%\MSC[2010] 00-01\sep  99-00
%\end{keyword}

@@

@@

\paragraph{Supermatrix}

1: Downloads GenBank database of the specified GB division (PLN, MAM, etc)
2: Perform exhaustive all-by-all BLAST comparisons of each ingroup and outgroup sequence.
3: Alternatively (much faster) uses a FASTA file of guide sequences to define each cluster.
   Each ingroup/outgroup sequence is BLASTed against the guide sequences.
   4: Build clusters of sequences:
           - use single-linkage hierarchical clustering algorithm
	           - distance threshold default BLASTn e-value 1.0e-10
		           - uses sequence length percent similarity cutoff default 0.5
			           - discards clusters that are not phylogenetically informative (< 4 taxa)
				   5: Aligns each cluster of sequences using MUSCLE.
				   6: Concatenates the clusters creating a supermatrix.

\paragraph{Phylogenetic analysis} Once the package is properly installed, you can use the document class \emph{elsarticle} to create a manuscript. Please make sure that your manuscript follows the guidelines in the Guide for Authors of the relevant journal. It is not necessary to typeset your manuscript in exactly the same way as an article, unless you are submitting to a camera-ready copy (CRC) journal.

Here is a table of the fossil calibrations used....

\paragraph{Chromosome evolution} The Elsevier article class is based on the standard article class and supports almost all of the functionality of that class. In addition, it features commands and options to format the
\begin{itemize}
\item document style
\item baselineskip
\item front matter
\item keywords and MSC codes
\item theorems, definitions and proofs
\item lables of enumerations
\item citation style and labeling.
\end{itemize}

@@

Here is a table with the mean divergence time estimates including \%95 confidence intervals....

The author names and affiliations could be formatted in two ways:
\begin{enumerate}[(1)]
\item Group the authors per affiliation.
\item Use footnotes to indicate the affiliations.
\end{enumerate}
See the front matter of this document for examples. You are recommended to conform your choice to the journal you are submitting to.

@@

@@

There are various bibliography styles availablei, please see Figure \ref{genera}. You can select the style of your choice in the preamble of this document. These styles are Elsevier styles based on standard styles like Harvard and Vancouver. Please use Bib\TeX\ to generate your bibliography and include DOIs whenever available.

Here are two sample references: \citep{Feynman1963118,Dirac1953888}.

>>>>

\usepackage{lineno}
\usepackage[hidelinks]{hyperref}
\usepackage{mathtools}
\usepackage{adjustbox}

\modulolinenumbers[5]

@@

%% bibliography styles
%% Harvard
\bibliographystyle{model2-names}\biboptions{authoryear}
%% `Elsevier LaTeX' style
%\bibliographystyle{elsarticle-num}

%%%%%%%%%%%%%%%%%%%%%%%
%% front matter
%%%%%%%%%%%%%%%%%%%%%%%

@@

\begin{abstract}
blah blah
\end{abstract}

@@

%%%%%%%%%%%%%%%%%%%%%%%
%% intro
%%%%%%%%%%%%%%%%%%%%%%%

@@

\paragraph{Supermatrix assembly}

I downloaded all available DNA sequences from GenBank release 200 PLN division and
performed an exhaustive all-by-all BLASTn \citep{blast} comparison for sequences in Onagraceae
and Lythraceae.
Using a BLASTn e-value of $1.0e^{-10}$ threshold and a sequence length
percent similarity cutoff of 0.5,
I constructed clusters of putative homologs using a single-linkage hierarchical clustering algorithm.
Subspecies names were removed from all sequences, and all but one sequence of each species was pruned from each cluster.
Clusters that were not phylogenetically informative ($< 4$ taxa) were discarded,
and each cluster was aligned using MUSCLE \citep{edgar2004muscle}. 
The alignments were concatenated by species, and any species that was not present in at least
two clusters was removed from the supermatrix.
Code written to assemble the supermatrix is availabe as the Python module 
Supermatrix Constructor (SUMAC) \citep[\url{https://github.com/wf8/sumac}]{sumac}.

\paragraph{Phylogenetic analyses} 
Maximum likelihood analyses were performed with RAxML-HPC \citep{raxml} on the CIPRES Scientific Gateway \citep{cipres} 
using the rapid bootstrap heuristic and the GTRCAT nucleotide substitution model.
I used the ML tree to select 15 species phylogenetically widely distributed in Lythraceae to act as outgroup; 
all other members of Lythraceae were subsequently removed from the supermatrix before estimating divergence times.
Bayesian estimates of divergence times were inferred using BEAST \citep{beast, beast2} on CIPRES and calibrated with five fossils 
identified with morphological synapomorphies (Table \ref{fossils}).
The \textit{Ludwigia} fossil pollen was dated broadly to the Paleocene \citep{grimsson}, so I set the prior to a normal distribution with a wide 
standard deviation to cover the entire time period.
For all other calibration points I used a lognormal prior distribution with the offset (the minimum age of the node) corresponding to the fossil age.
The BEAST analysis utilized the GTR+$\Gamma$ nucleotide substitution model with a relaxed molecular clock (uncorrelated lognormal model)
and a Yule process tree prior.
The Markov Chain Monte Carlo (MCMC) was run for 100 million generations, sampling every 10 thousand generations.
Tracer v1.6 \citep{tracer} was used to assess the MCMC output for parameter convergence and ensure that the effective sample size for all parameters was above 200.
The first 1000 trees were discarded as burn-in, and the remaining 9000 trees were summarized as a maximum clade credibility tree with mean divergence times.

\paragraph{Character evolution}
blah blah

\begin{table*}
   \begin{adjustbox}{max width=\textwidth}
      \begin{tabular}{lllllll}
         \hline
         Group & Age (Mya) & Prior Distribution & Mean & SD & Offset & Reference \\ \hline
         \textit{Circaea} (Onagraceae) & 12 & lognormal & 0.0 & 2.0 & 12 & \citep{grimsson} \\
         \textit{Epilobium} (Onagraceae) & 12 & lognormal & 0.0 & 2.0 & 12 & \citep{grimsson} \\
         S. Pacific \textit{Fuschia} (Onagraceae) & 23 & lognormal & 0.0 & 1.0 & 23 & \citep{lee2013fossil} \\
         \textit{Ludwigia} (Onagraceae) & Paleocene & normal & 60.0 & 3.0 & - & \citep{zhi} \\
         Lythraceae & 82 & lognormal & 0.0 & 2.0 & 82 & \citep{graham} \\
         \hline
      \end{tabular}
   \end{adjustbox}
   \caption{Fossil priors used in BEAST analyses.}
   \label{fossils}
\end{table*}

%%%%%%%%%%%%%%%%%%%%%%%
%% results
%%%%%%%%%%%%%%%%%%%%%%%

@@

\paragraph{Supermatrix assembly}
blah blah

\paragraph{Divergence time estimates}
Here is a table with the mean divergence time estimates including \%95 confidence intervals....

\paragraph{Character evolution}
blah blah

@@

%%%%%%%%%%%%%%%%%%%%%%%
%% conclusion
%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%
%% references
%%%%%%%%%%%%%%%%%%%%%%%
